Placing the spotted T Tauri star LkCa 4 on an HR diagram

### 2016-07-25 document/ms.tex
Overhaul the results from IGRINS data and clean up some errors in the text.

In this work we constrain the starspot properties of \name with 4 complementary techniques.  Section 2 and Appendix \ref{methods-details} describe a spectral inference technique aimed at generalizing line depth ratio analysis.  Section 3 applies the spectral inference technique to a high resolution panchromatic near-IR spectrum of \name.  Sections 4.1, 4.2, and 4.3 show results from polychromatic photometric monitoring, SED fitting, and optical TiO line-depth ratio fitting.  All lines of evidence are ultimately combined to build a consistent picture of the spectral and temporal evolution of \name, and what can be understood about pre-MS stellar evolution from this exemplar.

@@

The stellar photosphere is characterized as two photospheric components with a starspot temperature $\teffb$ and ambient photosphere temperature $\teffa$, with scalar solid angular coverages $\Omega_b$ and $\Omega_a$, respectively.    Starspots (or spot groups) are assumed to be cooler than the ambient photosphere, and they share the same average intrinisic and extrinsic stellar parameters otherwise; that is $(\vsini, \logg, \Z, v_z)$ are indentical for starspots and ambient photosphere.  The composite mixture model for observed flux density can be written down as:

@@

The starspot model includes two more free parameters than the standard \iancze\ model, namely $\teffb$ and $c$.  The addition of these two parameters makes the MCMC sampling much more correlated than it was before because the relative contribution of the starspot and ambient photosphere are nearly degnerate over small changes in effective temperature.  So the second important change to the spectral inference framework involves technical aspects of switching from sampling the nuisance and stellar parameters separately in a blocked Gibbs framework to ensemble sampling using \texttt{emcee} \citep{foreman13}.  The affine-invariant \texttt{emcee} ensemble sampler is more resilient to correlations among stellar temperatures and fill factor than the Metropolis-Hastings sampler used in Gibbs sampling.  The practical effect of this switch is that we obtain unique stellar parameter estimates $\vT_{o}$ for each spectral order $o$, whereas the \iancze\ strategy had the power to provide a single set of stellar parameters $\vT$ that was based on all $N_{\rm ord}$ spectral orders.  We combine the $N_{\rm ord}$ sets of inferences on $\vT_{o}$ with heuristics like taking the weighted average of point estimates of parameter distributions of the most reliable subsets of orders.  These lossy heuristics should regress to the mean with a large number of spectral orders, assuming each order delivers unbiased estimates of stellar parameters.

@@

Provide overview of section...

@@

The heightened sensitivity to starspot spectral lines as a function of wavelength can be understood in the following way.  The effective temperature of a starspot is, by definition, cooler than its surrounding photosphere, and will therefore have a longer wavelength of peak emission, according to Wein's displacement law.  So the ratio of flux density between starspot and bulk photosphere as a function of wavelength will generally increase with wavelength until it assymptotes to a fixed value in the Rayleigh-Jeans tail \citep{wolk96}.

%In Figure XX we plot the ratio of black body flux density at temperatures 4100 K and 3300 K with a red line.  The blue line is the ratio of two smoothed \PHOENIX model spectra with 4100 K and 3300 K, solar metalicity, and $\logg=3.5$.  The flux density ratio is the flux per solid angle, so the delivered flux ratio for a finite sized starspot will depend on the (flux-weighted) solid angle of the starspot relative to the (flux-weighted) solid angle of the stellar disk, vis-a-vis Equations \ref{eqn:Mix_M} and \ref{eqn:fill_factor}.  
%\todo{Put in the flux density ratio plot.}

\comm{Should we change the plot numbers from 4100 and 3300 to 4200 and 3000?  -These would end up being closer to the best fit values.}

In summary, the visible portion of the spectrum will be dominated by the warm patches, but the contrast with the starspots will be higher.  The infrared spectrum will have a lower overall spot contast, but will have higher relative contribution attributable to the cool patches.  This observation has motivated precision radial velocity (RV) planet searches in the infrared, where the modulation of RV signals is lessened \citep{prato08}.
\comm{This blurb should probably go in the introduction.}

@@

Motivated by the failure of single-temperature spectral fits, we added two new model parameters to our forward model: a starspot temperature and a fill factor, as described by Equation \ref{eqn:Mix_M}.  For computational complexity reasons, we altered the model framework as described in Section \ref{sec:methods} and detailed in Appendix \ref{methods-details}.

We computed full-spectrum fitting of 43 of 54 IGRINS spectral orders.  We fit all 14 stellar and nuissance parameters simultaneously.  We ran the fits for 5000 steps with 40 walkers in \texttt{emcee}.  The stellar parameter ranges were $\logg \in [-3.5, 4.0]$ and $\Z \in [-0.5, 0.5]$; the effective temperature range for $H-$ band was $\teffa, \teffb \in [3000, 4300]$, and was expanded to $\teffa, \teffb \in [2700, 4500]$ for $K-$band.  We used the same priors for stellar and nuissance parameters as \iancze.  We enforced $\teffb < \teffa$ to limit the combinatoric degeneracy and mode-mixing that's possible if MCMC samples with $\teffa \approx \teffb$ arise.  This false mode mixing matters for the \texttt{emcee} algorithm since the step size distribution would be artificially inflated if some walkers mixed modes with $\teffa \Leftrightarrow \teffb$.  The MCMC chains all appeared to converge after about 1500 steps.  \footnote{We spot-checked a few of the MCMC chains for convergence by computing the integrated autocorrelation times.  Most of the parameters had integrated autocorrelation times much less than our number of steps, except for the $\log{\Omega}$ terms.  This makes sense because the $\log{\Omega_1}$ and $\log{\Omega_2}$ are strongly correlated, so it takes many steps for each walker to explore the joint probability space.  We have a strong constraint on the filling factor with all 40 walkers combined.}

We computed point estimates for each parameter by selecting the final 200 $\times$ 40 walkers = 8000 samples, and computing the median, $5^{th}$ and $95^{th}$ percentiles of the marginal distributions.

We assessed the fit quality by examining the consistency of the point estimates across the spectral orders.  The distribution of $v_z$ and $\vsini$ were informative for finding extremely poor fits.  We find median point estimates and standard deviations of all 43 orders of $v_z = 12.4 \pm 2.6$ and $\vsini = 28.8 \pm 2.0$.  Spectral orders $m=91 \mathrm{ and }94$ had exceptionally large uncertainties on their values for $v_z$ and $\vsini$, which makes sense since these two orders are badly affected by telluric absorption artifacts.

As noted previously, our full-spectrum fitting method offers little constraining power on $\logg$, and $\Z$ can be biased in the face of systematic spectral line outliers.  We find a preference for values of $\logg$ on the high end of our search range, $3.5 <\logg <4.0$, finding a median value of the point estimates of 3.8 and a standard deviation of 0.1, consistent with previously published values from \citet{donati14}.  For $\Z$ we find a slight trend with wavelength-- the longest wavelength orders tend to prefer higher metalicity, while the shorter wavelength values prefer slightly sub-solar metalicity.  Orders 79, 81, 90, and to a lesser extent 78 and 76 showed aritificially tight constraints on $\Z$, biased to large values.  Similarly, orders 86, 101, 102, 104, 106, all had $\Z$ estimates biased to extremely low values.   Visual inspection of these orders indicates that their spectrum is dominated by spectral line outliers--deep metal lines that cannot be described by a coarse metalicity estimate alone.  These spectral line outlier orders also tend to bias estimates of effective temperature.  The interpretation of these lines will be addressed in the discussion section.

Finally, the posteriors on $\teffa$ and $\teffb$ show several trends- see Figure \ref{fig:TwoTempResults} and Table \ref{tbl_order_results} for the results of effective temperatures for all IGRINS orders.  At first sight, the $H-$ band estimates show remarkably tight distributions centered near $\teffa\sim4100$ K and $\teffb\sim3000$ K, with the exception of a few spectral order outliers- $m=$ 107, 108, 109. The apparent tight constraint is an illusion from pileup near our effective temperature search range for $H-$band.  So the point estimates for $\teffa$ and $\teffb$ should really be considered lower- and upper- limits respectively.  Orders 107, 108, and 109 converged on solutions with nearly equal effective tempatures for starspot and photosphere, $\teffa \sim \teffb$.  These nearly-equal solutions could be explained by vanishingly weak spectral lines in one or the other of the effective temperature ranges, in which case the method offers little constraint on $\teffb$, but improves the fit around $\teffa$.

We experimented on the $K-$band spectra with expanded effective temperature search ranges.  Here we see much more scatter, with no single photospheric temperature sticking out as strongly favored, but with values of $\teffa \sim 3900$ K and $\teffb \sim 2800$ K consistent with many orders.  The $K-$band spectra possess strongly correlated residuals from poor telluric correction, as well as uncorrected hydrogen line features that appear as flux excessess after division by telluric standard spectrum.  These instrumental afflictions degraded fits to several orders.

@@

Figures \ref{fig:Hband3x7} and \ref{fig:Kband3x7} in the Appendix show 42 of the $H-$ and $K-$band spectra on a log scale with a single random composite model spectrum overplotted.  The red and blue lines show the hot photosphere and cool spot spectra respectively.  We inspected the quality of the fits, generally finding remarkable agreement with most spectral orders.  Many orders demonstrated spectral line outliers, as noted in \iancze.  Most of the gross spectral features are captured by the composite model.  There are many examples of spectral lines that are present only in the hot photosphere model, or only in the cool photosphere model.

%% Examples of selected regions.
For example...

\todo{Compile a list of Line IDs, and maybe little postage stamp figures for each line with a range of fits.}

The most compelling results from the full spectrum fitting method come from orders for which there exist strongly temperature-sensitive spectral lines.  We spot-checked each spectrum, flagging the orders that possessed these types of lines.  We then computed the weighted mean of the stellar parameter point statistics only for these effective temperatures and fill factors.  We arrived at best fit values for the stellar parameters summarized in Table \ref{tbl_adopted_props}.  We adopt these parameters as the stellar properties at the epoch of the IGRINS observations.  We find a large filling factor of cool starspots, with a best value of $f=75\pm 10 \% $.

>>>>

In this work we constrain the starspot properties of \name with 4 complementary techniques.  Section 2 and Appendix \ref{methods-details} describe a spectral inference technique aimed at generalizing line depth ratio analysis.  Section 3 applies the spectral inference technique to a high resolution panchromatic near-IR spectrum of \name.  Sections 5.1, 5.2, and 5.3 show results from SED fitting, polychromatic photometric monitoring, and optical TiO line-depth ratio fitting.  All lines of evidence are ultimately combined to build a consistent picture of the spectral and temporal evolution of \name, and what can be understood about pre-MS stellar evolution from this exemplar.

@@

The stellar photosphere is characterized as two photospheric components with a starspot temperature $\teffb$ and ambient photosphere temperature $\teffa$, with scalar solid angular coverages $\Omega_b$ and $\Omega_a$, respectively.    Starspots (or spot groups) are assumed to be cooler than the ambient photosphere, and they share the same average intrinsic and extrinsic stellar parameters otherwise; that is $(\vsini, \logg, \Z, v_z)$ are identical for starspots and ambient photosphere.  The composite mixture model for observed flux density can be written down as:

@@

The starspot model includes two more free parameters than the standard \iancze\ model, namely $\teffb$ and $c$.  The addition of these two parameters makes the MCMC sampling much more correlated than it was before because the relative contribution of the starspot and ambient photosphere are nearly degenerate over small changes in effective temperature.  So the second important change to the spectral inference framework involves technical aspects of switching from sampling the nuisance and stellar parameters separately in a blocked Gibbs framework to ensemble sampling using \texttt{emcee} \citep{foreman13}.  The affine-invariant \texttt{emcee} ensemble sampler is more resilient to correlations among stellar temperatures and fill factor than the Metropolis-Hastings sampler used in Gibbs sampling.  The practical effect of this switch is that all 14 stellar and nuisance parameters are fit simultaneously in a single spectral order.  So stellar parameter estimates $\vT_{o}$ are unique for each spectral order $o$, whereas the \iancze\ strategy had the power to provide a single set of stellar parameters $\vT$ that was based on all $N_{\rm ord}$ spectral orders.  The $N_{\rm ord}$ sets of inferences on $\vT_{o}$ are combined by weighted averaging point estimates of results from reliable orders, which offers resilience to flagrant calibration artifacts or clear model mismatches.  This process can be thought of as a course approximation to the much more sophisticated spectral line outlier rejection described in \iancze.

@@

\todo{Provide overview of section...}

@@

The heightened sensitivity to starspot spectral lines as a function of wavelength can be understood in the following way.  Starspots are cooler than their surrounding photosphere and will therefore have a longer wavelength of peak emission, according to Wein's displacement law.  The average ratio of flux density between starspot and bulk photosphere will increase with wavelength until asymptoting to a fixed value in the Rayleigh-Jeans tail \citep{wolk96}.

%In Figure XX we plot the ratio of black body flux density at temperatures 4100 K and 3300 K with a red line.  The blue line is the ratio of two smoothed \PHOENIX model spectra with 4100 K and 3300 K, solar metalicity, and $\logg=3.5$.  The flux density ratio is the flux per solid angle, so the delivered flux ratio for a finite sized starspot will depend on the (flux-weighted) solid angle of the starspot relative to the (flux-weighted) solid angle of the stellar disk, vis-a-vis Equations \ref{eqn:Mix_M} and \ref{eqn:fill_factor}.  
%\todo{Put in the flux density ratio plot.}
%\comm{Should we change the plot numbers from 4100 and 3300 to 4200 and 3000?  -These would end up being closer to the best fit values.}

In summary, the visible portion of the spectrum will be dominated by the warm patches, but the contrast with the starspots will be higher.  The infrared spectrum will have a lower overall spot contast, but will have higher relative contribution attributable to the cool patches.  This observation has motivated precision radial velocity (RV) planet searches in the infrared, where the modulation of RV signals is lessened \citep{prato08}.
%\comm{This blurb should probably go in the introduction.}

@@

Full-spectrum fitting was performed for 43 of the 54 available IGRINS spectral orders as described in Sectrion \ref{sec:methods} and Appendix \ref{methods-details}.  The MCMC samples had 5000 steps with 40 walkers.  The stellar parameter ranges were $\logg \in [-3.5, 4.0]$ and $\Z \in [-0.5, 0.5]$; the effective temperature range for $H-$ band was $\teffa, \teffb \in [2700, 4500]$, and was expanded to $\teffa, \teffb \in [2700, 4500]$ for $K-$band.  The MCMC chains all appeared to converge after about 1500 steps.

Marginalized point estimates were computed for all 14 stellar and nuisance parameters by selecting the final 200 $\times$ 40 walkers = 8000 samples, and computing the median, $5^{th}$ and $95^{th}$ percentiles of the marginal distributions.  The fit quality was assessed by examining the consistency of the point estimates across the spectral orders.  The distribution of $v_z$ and $\vsini$ were informative for finding extremely poor fits.  The range for all 43 orders was $v_z = 12.4 \pm 2.6$ and $\vsini = 28.8 \pm 2.0$.  Spectral orders $m=91$ and $94$ had exceptionally large uncertainties on their values for $v_z$ and $\vsini$, which makes sense since these two orders are badly affected by telluric absorption artifacts.

As noted previously, the full-spectrum fitting method offers little constraining power on $\logg$, and $\Z$ can be biased in the face of systematic spectral line outliers.  Estimates of $\logg$ appeared the high end of our search range, $3.5 <\logg <4.0$, yielding a median value of the point estimates of 3.8 and a standard deviation of 0.1, consistent with previously published values from \citet{donati14}.  The estimates for $\Z$ yield a slight trend with wavelength-- the longest wavelength orders tend to prefer higher metalicity, while the shorter wavelength values prefer slightly sub-solar metalicity.  Orders 79, 81, 90, and to a lesser extent 78 and 76 showed aritificially tight constraints on $\Z$, biased to large values.  Similarly, orders 86, 101, 102, 104, 106, all had $\Z$ estimates biased to extremely low values.   Visual inspection of these orders indicates that their spectrum is dominated by spectral line outliers--deep metal lines that cannot be described by a coarse metalicity estimate alone.  These spectral line outlier orders also tend to bias estimates of effective temperature.  These metal lines can also be biased by Zeeman broadening, which would explain the heightened effect size with wavelength.

Finally, the posteriors on $\teffa$ and $\teffb$ show several trends- see Figure \ref{fig:TwoTempResults} and Table \ref{tbl_order_results} for the results of effective temperatures for all IGRINS orders.  
\todo{insert results from re-run of H-band spectra here}
Nearly-equal solutions could be explained by vanishingly weak spectral lines in one or the other of the effective temperature ranges, in which case the method offers little constraint on $\teffb$, but improves the fit around $\teffa$.

The $K-$band orders demonstrate much more scatter than $H-$band, with no single photospheric temperature sticking out as strongly favored, but with values of $\teffa \sim 3900$ K and $\teffb \sim 2800$ K consistent with many orders.  The $K-$band spectra possess strongly correlated residuals from poor telluric correction, as well as uncorrected hydrogen line features that appear as flux excessess after division by telluric standard spectrum.  These instrumental afflictions degraded fits to several orders.

@@

Figures \ref{fig:Hband3x7} and \ref{fig:Kband3x7} in the Appendix show 42 of the $H-$ and $K-$band spectra on a log scale with a single random composite model spectrum overplotted.  The red and blue lines show the hot photosphere and cool spot spectra respectively.  We inspected the quality of the fits, generally finding remarkable agreement with most spectral orders.  Many orders demonstrated spectral line outliers, as noted in \iancze.  Most of the gross spectral features are captured by the composite model.  There are many examples of spectral lines that are present only in the hot photosphere model, or only in the cool photosphere model.  For example, the OH 1.563 $\mu$m line noted by \citet{oneal01} shows a clear pattern of 3 lines in the data, with the central line exceeding the depth of the adjacent two lines.

Best fit values for the stellar parameters summarized are listen in Table \ref{tbl_adopted_props}.  Remarkably, the filling factor of starspots exceeds 50\% with a best fit value of $c=75\pm 10 \% $.
